---
title: "Pre-processing"
author: "Nick Stern"
date: "4/28/2019"
output: 
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    fig_width: 6
    fig_height: 4
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This markdown files cleans the Acute Myocardial Infarction (AMI) dataset, taken from the Canvas course page. The step by step pre-processing actions are outlined below.

First we read in the dataset

```{r message=FALSE, warning=FALSE}
## Reading in the original data
library(dplyr)
ami = read.csv('data/amidata.csv')
head(ami)
```

Then we made the variable names lowercase for ease of use and to adhere to convention.

```{r}
# make variable names lowercase
ami.names<-tolower(colnames(ami))
colnames(ami)<-ami.names
```

Then we imputed the missing values in the ```charges``` column with the ```na.convert.mean``` function from lecture notes. This function imputes missing values with the mean of the predictor, and creates a new binary predictor that is essentially an indicator variable for which values were originally missing. 

```{r}
source('na-convert.R')  # NA conversion function from class
# impute missing values
ami <- na.convert.mean(ami)
```

Finally, we factorized some of the categorical variables that weren't factorized by default.

```{r}
#recode variables
ami <- ami %>%
  mutate(diagnosis = factor(diagnosis),
         drg = factor(drg),
         died = factor(died))
```

Below we provide a summary of the pre-processed dataset, and save it to the local filesystem.

```{r}
#summary statistics
summary(ami)
saveRDS(ami, file='data/amidata_clean.rds')
```




