---
title: "Stat149: Final Project Report"
author: "Karina Huang, Nicholas Stern, Phoebe Wong"
date: "4/18/2019"
output: 
  pdf_document:
    number_sections: true
---

```{r imports, include=FALSE}
library(tidyverse)
```


# Background
- Description of the problem and its importance
- Description of data
- Goals of project

# Exploratory Analyses
- Key features of the data that will be relevant for modeling
- Any data preprocessing?

```{r read-in-data, echo=FALSE, warning=FALSE, message=FALSE}
data <- read_csv("data/amidata.csv")
```

## EDA 
### Response variable - Hospital legnth of stay

```{r eda-response, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(data, aes(LOS)) +
  geom_histogram() + 
  labs(x = "Length of stay (Days)", y = "Count", 
       title = "Histogram of hospital length of stay") +
  theme_minimal()
```

# Model
- Main models considered
- Decisions to consider alternative models

# Conclusion
- Substantive conclusions that would be important for a non-statistican
- Report any prgamatic advice/conclusions result from the analysis

# Discussion
- Critical evaluation of your overall approach
- What aspects of the modeling attempts did you expect would substantially improve inference but did not achieve the desired outcome?
- What limitations can you offer about your results or the process that led to your results?